---
title: "Augmentation Methods: Performance Comparison"
author: "Oktawia Miluch"
date: today
bibliography: ../readme_files/references.bib
link-citations: true
format:
  html:
    toc: true
    number-sections: true
    theme: cosmo
    fig-format: png
    code-fold: true
    embed-resources: true
execute:
  echo: true
  warning: false
  message: false
knitr:
  opts_chunk:
    fig.path: _figures/methods_comparison/
    fig.width: 18
    fig.height: 13.5
---

```{r}
#| label: setup
#| include: false
library(here)
library(purrr)
library(dplyr)
library(tidyr)
library(tibble)
library(stringr)
library(readr)
library(ggplot2)
library(viridis)
library(crosstalk)
library(DT)
library(htmltools)
library(kableExtra)

# Create/clean graphics directory, set fig.path
graphics_abs <- here::here("content", "_figures", "methods_comparison")
dir.create(graphics_abs, recursive = TRUE, showWarnings = FALSE)

# clean contents but keep folder
to_remove <- list.files(
  graphics_abs,
  all.files = TRUE,
  full.names = TRUE,
  no.. = TRUE
)
unlink(to_remove, recursive = TRUE, force = TRUE)

# ensure knitr writes to the selected path
knitr::opts_chunk$set(
  fig.path   = "_figures/methods_comparison/",
  fig.width  = 18,
  fig.height = 13.5
)

# Source helper functions
source(file.path(here::here(), "R", "methods_comparison_helpers.R"))
source(file.path(here::here(), "R", "summary_rodriguez_helpers.R"))
source(file.path(here::here(), "R", "summary_quarto_docs_helpers.R"))

# Read in aggregated results from Rodriguez augmentation methods
data_dir <- here::here("results", "results_aggregated")
## pc = 1/max_lib_size
rod_files <- list.files(
  data_dir,
  pattern = "^rodriguez.*pc_max_lib_size$",
  full.names = TRUE
)

rod_names <- rod_files |>
  basename()

rod_data <- map(rod_files, readRDS)

walk2(rod_data, rod_names, ~ assign(.y, .x, envir = .GlobalEnv))
rm(rod_data)

## pc = 1/2
rod_files_half <- list.files(
  data_dir,
  pattern = "^rodriguez.*pc_half$",
  full.names = TRUE
)

rod_names_half <- rod_files_half |>
  basename()

rod_data_half <- map(rod_files_half, readRDS)

walk2(rod_data_half, rod_names_half, ~ assign(.y, .x, envir = .GlobalEnv))
rm(rod_data_half)

# Read in aggregated results from random ILR augmentation methods
## PC = 1/ max lib size
ilr_files <- list.files(
  data_dir,
  pattern = "^random_ilr.*pc_max_lib_size$",
  full.names = TRUE
)

ilr_names <- ilr_files |>
  basename()

ilr_data <- map(ilr_files, readRDS)

walk2(ilr_data, ilr_names, ~ assign(.y, .x, envir = .GlobalEnv))
rm(ilr_data)

## PC = 1/2
ilr_files_half <- list.files(
  data_dir,
  pattern = "^random_ilr.*pc_half$",
  full.names = TRUE
)

ilr_names_half <- ilr_files_half |>
  basename()

ilr_data_half <- map(ilr_files_half, readRDS)

walk2(ilr_data_half, ilr_names_half, ~ assign(.y, .x, envir = .GlobalEnv))
rm(ilr_data_half)

# Benchmark
## PC = 1/ max lib size
bench_files <- list.files(
  data_dir,
  pattern = "^benchmark.*pc_max_lib_size$",
  full.names = TRUE
)

bench_names <- bench_files |>
  basename()

bench_data <- map(bench_files, readRDS)

walk2(bench_data, bench_names, ~ assign(.y, .x, envir = .GlobalEnv))
rm(bench_data)

## PC = 1/2
bench_files_half <- list.files(
  data_dir,
  pattern = "^benchmark.*pc_half$",
  full.names = TRUE
)

bench_names_half <- bench_files_half |>
  basename()

bench_data_half <- map(bench_files_half, readRDS)

walk2(bench_data_half, bench_names_half, ~ assign(.y, .x, envir = .GlobalEnv))
rm(bench_data_half)
```

# The purpose of this document

We compare the performance achieved by each model across the considered augmentation techniques and their benchmarks.

Plan for this document:
1. create tables summarizing the mean performance across augmentation techniques
2. create boxplots showing the distribution of performance metrics
3. plot the relation between n/p ratio and performance
4. build a heatmap to check whether the model outperforms (i) its benchmark and (ii) the standard ILR benchmark

# Logistic Regression with L1 Penalty

## ROC AUC
### Summary Table - Mean Performance Metric
```{r}
#| label: roc_auc_mean_lasso
#| echo: false
#| results: asis
render_mean_metric_sections(
    rand_ilr_half = random_ilr_lasso_pc_half,
    rand_ilr_max = random_ilr_lasso_pc_max_lib_size,
    benchmark_max = benchmark_lasso_pc_max_lib_size,
    benchmark_half = benchmark_lasso_pc_half,
    rod_half = rodriguez_lasso_pc_half,
    rod_max = rodriguez_lasso_pc_max_lib_size,
    plot_metric = "roc_auc"
)
```

### Boxplots
::: panel-tabset
```{r}
#| results: asis
#| label: auc_boxplot_lasso
render_metric_facets_tabset(
  rand_ilr_half = random_ilr_lasso_pc_half,
  rand_ilr_max = random_ilr_lasso_pc_max_lib_size,
  benchmark_max = benchmark_lasso_pc_max_lib_size,
  benchmark_half = benchmark_lasso_pc_half,
  rod_half = rodriguez_lasso_pc_half,
  rod_max = rodriguez_lasso_pc_max_lib_size,
  plot_metric = "roc_auc"
)
```
:::


```{r}
#| label: n_p_of_datasets
#| echo: false
# Capture dimensionality (p) and sample size (n) for each preprocessed dataset
# The resulst of this code snippet are re-used in multiple 
directory_preproc <- here::here("data", "data_preproc")
preproc_files <- list.files(
  directory_preproc,
  pattern = "^task\\d+_preproc_x_pc_half\\.csv$",
  full.names = TRUE
)

data_dim <- preproc_files |>
  map(\(path) {
    dataset <- readr::read_csv(path, show_col_types = FALSE)
    feature_cols <- setdiff(names(dataset), "sample_id")

    tibble(
      data_id = as.integer(readr::parse_number(basename(path))),
      p = length(feature_cols),
      n = nrow(dataset)
    )
  })  |> 
  bind_rows() |>
  arrange(data_id)
```

### n/p Ratio

The plots below show the gain in ROC AUC from augmentation as a function of the original dataset's **n/p ratio before augmentation** (ratio between number of samples and number of features). Each tab corresponds to a single pseudo-count.

#### Pseudo-count 1/2
::: panel-tabset
```{r}
#| results: asis
#| label: n_p_plot_half_lasso_roc
render_aug_factor_tabset(
  rand_ilr_res = random_ilr_lasso_pc_half,
  rod_res = rodriguez_lasso_pc_half,
  benchmark = benchmark_lasso_pc_half,
  plot_metric = "roc_auc",
  data_dim = data_dim
)
```
:::

#### Pseudo-count 1 / max library size
::: panel-tabset
```{r}
#| results: asis
#| label: n_p_plot_max_lib_lasso_roc
render_aug_factor_tabset(
  rand_ilr_res = random_ilr_lasso_pc_max_lib_size,
  rod_res = rodriguez_lasso_pc_max_lib_size,
  benchmark = benchmark_lasso_pc_max_lib_size,
  plot_metric = "roc_auc",
  data_dim = data_dim
)
```
:::

## Misclassification Rate
### Summary Table - Mean Performance Metric
```{r}
#| label: misclas_mean_lasso
#| echo: false
#| results: asis
render_mean_metric_sections(
    rand_ilr_half = random_ilr_lasso_pc_half,
    rand_ilr_max = random_ilr_lasso_pc_max_lib_size,
    benchmark_max = benchmark_lasso_pc_max_lib_size,
    benchmark_half = benchmark_lasso_pc_half,
    rod_half = rodriguez_lasso_pc_half,
    rod_max = rodriguez_lasso_pc_max_lib_size,
    plot_metric = "misclassification_rate"
)
```

### Boxplots
::: panel-tabset
```{r}
#| results: asis
#| label: misclas_boxplot_lasso
render_metric_facets_tabset(
  rand_ilr_half = random_ilr_lasso_pc_half,
  rand_ilr_max = random_ilr_lasso_pc_max_lib_size,
  benchmark_max = benchmark_lasso_pc_max_lib_size,
  benchmark_half = benchmark_lasso_pc_half,
  rod_half = rodriguez_lasso_pc_half,
  rod_max = rodriguez_lasso_pc_max_lib_size,
  plot_metric = "misclassification_rate"
)
```
:::

### n/p Ratio

The plots below show the gain in misclassification rate from augmentation (computed as the difference to the benchmark) against the original dataset's **n/p ratio before augmentation**. Each tab corresponds to a single pseudo-count.

#### Pseudo-count 1/2
::: panel-tabset
```{r}
#| results: asis
#| label: n_p_plot_half_lasso_misclas
render_aug_factor_tabset(
  rand_ilr_res = random_ilr_lasso_pc_half,
  rod_res = rodriguez_lasso_pc_half,
  benchmark = benchmark_lasso_pc_half,
  plot_metric = "misclassification_rate",
  data_dim = data_dim
)
```
:::

#### Pseudo-count 1 / max library size
::: panel-tabset
```{r}
#| results: asis
#| label: n_p_plot_max_lib_lasso_misclas
render_aug_factor_tabset(
  rand_ilr_res = random_ilr_lasso_pc_max_lib_size,
  rod_res = rodriguez_lasso_pc_max_lib_size,
  benchmark = benchmark_lasso_pc_max_lib_size,
  plot_metric = "misclassification_rate",
  data_dim = data_dim
)
```
:::

# Random Forest 

## ROC AUC
### Summary Table - Mean Performance Metric
```{r}
#| label: roc_auc_mean_rf
#| echo: false
#| results: asis
render_mean_metric_sections(
    rand_ilr_half = random_ilr_random_forest_pc_half,
    rand_ilr_max = random_ilr_random_forest_pc_max_lib_size,
    benchmark_max = benchmark_random_forest_pc_max_lib_size,
    benchmark_half = benchmark_random_forest_pc_half,
    rod_half = rodriguez_random_forest_pc_half,
    rod_max = rodriguez_random_forest_pc_max_lib_size,
    plot_metric = "roc_auc"
)
```

### Boxplots
::: panel-tabset
```{r}
#| results: asis
#| label: auc_boxplot_rf
render_metric_facets_tabset(
  rand_ilr_half = random_ilr_random_forest_pc_half,
  rand_ilr_max = random_ilr_random_forest_pc_max_lib_size,
  benchmark_max = benchmark_random_forest_pc_max_lib_size,
  benchmark_half = benchmark_random_forest_pc_half,
  rod_half = rodriguez_random_forest_pc_half,
  rod_max = rodriguez_random_forest_pc_max_lib_size,
  plot_metric = "roc_auc"
)
```
:::

### n/p Ratio

The plots below show the gain in ROC AUC from augmentation as a function of the original dataset's **n/p ratio before augmentation**. Each tab corresponds to a single pseudo-count.

#### Pseudo-count 1/2
::: panel-tabset
```{r}
#| results: asis
#| label: n_p_plot_half_rf_roc
render_aug_factor_tabset(
  rand_ilr_res = random_ilr_random_forest_pc_half,
  rod_res = rodriguez_random_forest_pc_half,
  benchmark = benchmark_random_forest_pc_half,
  plot_metric = "roc_auc",
  data_dim = data_dim
)
```
:::

#### Pseudo-count 1 / max library size
::: panel-tabset
```{r}
#| results: asis
#| label: n_p_plot_max_lib_rf_roc
render_aug_factor_tabset(
  rand_ilr_res = random_ilr_random_forest_pc_max_lib_size,
  rod_res = rodriguez_random_forest_pc_max_lib_size,
  benchmark = benchmark_random_forest_pc_max_lib_size,
  plot_metric = "roc_auc",
  data_dim = data_dim
)
```
:::

## Misclassification Rate
### Summary Table - Mean Performance Metric
```{r}
#| label: misclas_mean_rf
#| echo: false
#| results: asis
render_mean_metric_sections(
    rand_ilr_half = random_ilr_random_forest_pc_half,
    rand_ilr_max = random_ilr_random_forest_pc_max_lib_size,
    benchmark_max = benchmark_random_forest_pc_max_lib_size,
    benchmark_half = benchmark_random_forest_pc_half,
    rod_half = rodriguez_random_forest_pc_half,
    rod_max = rodriguez_random_forest_pc_max_lib_size,
    plot_metric = "misclassification_rate"
)
```

### Boxplots
::: panel-tabset
```{r}
#| results: asis
#| label: misclas_boxplot_rf
render_metric_facets_tabset(
  rand_ilr_half = random_ilr_random_forest_pc_half,
  rand_ilr_max = random_ilr_random_forest_pc_max_lib_size,
  benchmark_max = benchmark_random_forest_pc_max_lib_size,
  benchmark_half = benchmark_random_forest_pc_half,
  rod_half = rodriguez_random_forest_pc_half,
  rod_max = rodriguez_random_forest_pc_max_lib_size,
  plot_metric = "misclassification_rate"
)
```
:::

### n/p Ratio

The plots below show the gain in misclassification rate from augmentation (difference relative to the benchmark) against the original dataset's **n/p ratio before augmentation**. Each tab corresponds to a single pseudo-count.

#### Pseudo-count 1/2
::: panel-tabset
```{r}
#| results: asis
#| label: n_p_plot_half_rf_misclas
render_aug_factor_tabset(
  rand_ilr_res = random_ilr_random_forest_pc_half,
  rod_res = rodriguez_random_forest_pc_half,
  benchmark = benchmark_random_forest_pc_half,
  plot_metric = "misclassification_rate",
  data_dim = data_dim
)
```
:::

#### Pseudo-count 1 / max library size
::: panel-tabset
```{r}
#| results: asis
#| label: n_p_plot_max_lib_rf_misclas
render_aug_factor_tabset(
  rand_ilr_res = random_ilr_random_forest_pc_max_lib_size,
  rod_res = rodriguez_random_forest_pc_max_lib_size,
  benchmark = benchmark_random_forest_pc_max_lib_size,
  plot_metric = "misclassification_rate",
  data_dim = data_dim
)
```
:::

# XGBoost

## ROC AUC
### Summary Table - Mean Performance Metric
```{r}
#| label: roc_auc_mean_xgb
#| echo: false
#| results: asis
render_mean_metric_sections(
    rand_ilr_half = random_ilr_xgboost_pc_half,
    rand_ilr_max = random_ilr_xgboost_pc_max_lib_size,
    benchmark_max = benchmark_xgboost_pc_max_lib_size,
    benchmark_half = benchmark_xgboost_pc_half,
    rod_half = rodriguez_xgboost_pc_half,
    rod_max = rodriguez_xgboost_pc_max_lib_size,
    plot_metric = "roc_auc"
)
```

### Boxplots
::: panel-tabset
```{r}
#| results: asis
#| label: auc_boxplot_xgb
render_metric_facets_tabset(
  rand_ilr_half = random_ilr_xgboost_pc_half,
  rand_ilr_max = random_ilr_xgboost_pc_max_lib_size,
  benchmark_max = benchmark_xgboost_pc_max_lib_size,
  benchmark_half = benchmark_xgboost_pc_half,
  rod_half = rodriguez_xgboost_pc_half,
  rod_max = rodriguez_xgboost_pc_max_lib_size,
  plot_metric = "roc_auc"
)
```
:::

### n/p Ratio

The plots below show the gain in ROC AUC from augmentation as a function of the original dataset's **n/p ratio before augmentation**. Each tab corresponds to a single pseudo-count.

#### Pseudo-count 1/2
::: panel-tabset
```{r}
#| results: asis
#| label: n_p_plot_half_xgb_roc
render_aug_factor_tabset(
  rand_ilr_res = random_ilr_xgboost_pc_half,
  rod_res = rodriguez_xgboost_pc_half,
  benchmark = benchmark_xgboost_pc_half,
  plot_metric = "roc_auc",
  data_dim = data_dim
)
```
:::

#### Pseudo-count 1 / max library size
::: panel-tabset
```{r}
#| results: asis
#| label: n_p_plot_max_lib_xgb_roc
render_aug_factor_tabset(
  rand_ilr_res = random_ilr_xgboost_pc_max_lib_size,
  rod_res = rodriguez_xgboost_pc_max_lib_size,
  benchmark = benchmark_xgboost_pc_max_lib_size,
  plot_metric = "roc_auc",
  data_dim = data_dim
)
```
:::

## Misclassification Rate
### Summary Table - Mean Performance Metric
```{r}
#| label: misclas_mean_xgb
#| echo: false
#| results: asis
render_mean_metric_sections(
    rand_ilr_half = random_ilr_xgboost_pc_half,
    rand_ilr_max = random_ilr_xgboost_pc_max_lib_size,
    benchmark_max = benchmark_xgboost_pc_max_lib_size,
    benchmark_half = benchmark_xgboost_pc_half,
    rod_half = rodriguez_xgboost_pc_half,
    rod_max = rodriguez_xgboost_pc_max_lib_size,
    plot_metric = "misclassification_rate"
)
```

### Boxplots
::: panel-tabset
```{r}
#| results: asis
#| label: misclas_boxplot_xgb
render_metric_facets_tabset(
  rand_ilr_half = random_ilr_xgboost_pc_half,
  rand_ilr_max = random_ilr_xgboost_pc_max_lib_size,
  benchmark_max = benchmark_xgboost_pc_max_lib_size,
  benchmark_half = benchmark_xgboost_pc_half,
  rod_half = rodriguez_xgboost_pc_half,
  rod_max = rodriguez_xgboost_pc_max_lib_size,
  plot_metric = "misclassification_rate"
)
```
:::

### n/p Ratio

The plots below show the gain in misclassification rate from augmentation (difference relative to the benchmark) against the original dataset's **n/p ratio before augmentation**. Each tab corresponds to a single pseudo-count.

#### Pseudo-count 1/2
::: panel-tabset
```{r}
#| results: asis
#| label: n_p_plot_half_xgb_misclas
render_aug_factor_tabset(
  rand_ilr_res = random_ilr_xgboost_pc_half,
  rod_res = rodriguez_xgboost_pc_half,
  benchmark = benchmark_xgboost_pc_half,
  plot_metric = "misclassification_rate",
  data_dim = data_dim
)
```
:::

#### Pseudo-count 1 / max library size
::: panel-tabset
```{r}
#| results: asis
#| label: n_p_plot_max_lib_xgb_misclas
render_aug_factor_tabset(
  rand_ilr_res = random_ilr_xgboost_pc_max_lib_size,
  rod_res = rodriguez_xgboost_pc_max_lib_size,
  benchmark = benchmark_xgboost_pc_max_lib_size,
  plot_metric = "misclassification_rate",
  data_dim = data_dim
)
```
:::
