---
title: "Impact of pseudo-count"
author: "Oktawia Miluch"
date: today
bibliography: ../readme_files/references.bib
link-citations: true
format:
  html:
    toc: true
    number-sections: true
    theme: cosmo
    fig-format: png
    code-fold: true
    embed-resources: true
execute:
  echo: true
  warning: false
  message: false
knitr:
  opts_chunk:
    fig.path: _figures/pseudo_count_impact/
    fig.width: 18
    fig.height: 13.5
---

```{r}
#| label: setup
#| include: false
library(here)
library(purrr)
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(viridis)
library(crosstalk)
library(DT)
library(htmltools)
library(kableExtra)

# Create/clean graphics directory, set fig.path
graphics_abs <- here::here("content", "_figures", "pseudo_count_impact")
dir.create(graphics_abs, recursive = TRUE, showWarnings = FALSE)

# clean contents but keep folder
to_remove <- list.files(
  graphics_abs,
  all.files = TRUE,
  full.names = TRUE,
  no.. = TRUE
)
unlink(to_remove, recursive = TRUE, force = TRUE)

# ensure knitr writes to the selected path
knitr::opts_chunk$set(
  fig.path   = "_figures/pseudo_count_impact/",
  fig.width  = 18,
  fig.height = 13.5
)

# Source helper functions
source(file.path(here::here(), "R", "pseudo_count_impact_helpers.R"))

# Read in aggregated results from Rodriguez augmentation methods
data_dir <- here::here("results", "results_aggregated")
## pc = 1/max_lib_size
rod_files <- list.files(
  data_dir,
  pattern = "^rodriguez.*pc_max_lib_size$",
  full.names = TRUE
)

rod_names <- rod_files |>
  basename()

rod_data <- map(rod_files, readRDS)

walk2(rod_data, rod_names, ~ assign(.y, .x, envir = .GlobalEnv))
rm(rod_data)

## pc = 1/2
rod_files_half <- list.files(
  data_dir,
  pattern = "^rodriguez.*pc_half$",
  full.names = TRUE
)

rod_names_half <- rod_files_half |>
  basename()

rod_data_half <- map(rod_files_half, readRDS)

walk2(rod_data_half, rod_names_half, ~ assign(.y, .x, envir = .GlobalEnv))
rm(rod_data_half)

# Read in aggregated results from random ILR augmentation methods
## PC = 1/ max lib size
ilr_files <- list.files(
  data_dir,
  pattern = "^random_ilr.*pc_max_lib_size$",
  full.names = TRUE
)

ilr_names <- ilr_files |>
  basename()

ilr_data <- map(ilr_files, readRDS)

walk2(ilr_data, ilr_names, ~ assign(.y, .x, envir = .GlobalEnv))
rm(ilr_data)

## PC = 1/2
ilr_files_half <- list.files(
  data_dir,
  pattern = "^random_ilr.*pc_half$",
  full.names = TRUE
)

ilr_names_half <- ilr_files_half |>
  basename()

ilr_data_half <- map(ilr_files_half, readRDS)

walk2(ilr_data_half, ilr_names_half, ~ assign(.y, .x, envir = .GlobalEnv))
rm(ilr_data_half)

# Benchmark
## PC = 1/ max lib size
bench_files <- list.files(
  data_dir,
  pattern = "^benchmark.*pc_max_lib_size$",
  full.names = TRUE
)

bench_names <- bench_files |>
  basename()

bench_data <- map(bench_files, readRDS)

walk2(bench_data, bench_names, ~ assign(.y, .x, envir = .GlobalEnv))
rm(bench_data)

## PC = 1/2
bench_files_half <- list.files(
  data_dir,
  pattern = "^benchmark.*pc_half$",
  full.names = TRUE
)

bench_names_half <- bench_files_half |>
  basename()

bench_data_half <- map(bench_files_half, readRDS)

walk2(bench_data_half, bench_names_half, ~ assign(.y, .x, envir = .GlobalEnv))
rm(bench_data_half)
```

The purpose of this document is to analyse how the pseudo-count affects model performance depending on the augmentation strategy and the augmentation factor. In all plots below, the phrase **mean difference of the selected performance metric** for a given augmentation strategy, augmentation factor and dataset refers to:

$$
\begin{aligned}
  \text{Mean difference}
    &= \frac{1}{20} \sum_{i=1}^{20} \Bigl(
         \text{performance metric}_{\text{pseudo-count}=0.5}(i) \\
    &\quad - \text{performance metric}_{\text{pseudo-count}=1/\max\{\text{library size}\}}(i)
       \Bigr)
\end{aligned}
$$

The mean of the performance metric is computed over $20$ train–test data splits. Below, each plot corresponds to one model (logistic regression with $\text{L}_1$ penalty, Random Forest, or XGBoost) and one performance metric (ROC AUC or misclassification rate). Each tabset corresponds to a selected augmentation factor.

# Logistic regression with $\text{L}_1$ penalty 
## ROC AUC

::: panel-tabset
```{r}
#| label: lasso_roc_auc
#| echo: false
#| results: asis
render_mean_diff_heatmap_tabs(
  rand_ilr_half = random_ilr_lasso_pc_half,
  rand_ilr_max = random_ilr_lasso_pc_max_lib_size,
  rod_half = rodriguez_lasso_pc_half,
  rod_max = rodriguez_lasso_pc_max_lib_size,
  bench_half = benchmark_lasso_pc_half,
  bench_max = benchmark_lasso_pc_max_lib_size,
  plot_metric = "roc_auc"
)
```
:::

## Misclassification Rate

::: panel-tabset
```{r}
#| label: lasso_misclas_rate
#| echo: false
#| results: asis
render_mean_diff_heatmap_tabs(
  rand_ilr_half = random_ilr_lasso_pc_half,
  rand_ilr_max = random_ilr_lasso_pc_max_lib_size,
  rod_half = rodriguez_lasso_pc_half,
  rod_max = rodriguez_lasso_pc_max_lib_size,
  bench_half = benchmark_lasso_pc_half,
  bench_max = benchmark_lasso_pc_max_lib_size,
  plot_metric = "misclassification_rate"
)
```
:::

# Random Forest
## ROC AUC

::: panel-tabset
```{r}
#| label: rf_roc_auc
#| echo: false
#| results: asis
render_mean_diff_heatmap_tabs(
  rand_ilr_half = random_ilr_random_forest_pc_half,
  rand_ilr_max = random_ilr_random_forest_pc_max_lib_size,
  rod_half = rodriguez_random_forest_pc_half,
  rod_max = rodriguez_random_forest_pc_max_lib_size,
  bench_half = benchmark_random_forest_pc_half,
  bench_max = benchmark_random_forest_pc_max_lib_size,
  plot_metric = "roc_auc"
)
```
:::

## Misclassification Rate
::: panel-tabset
```{r}
#| label: rf_misclas
#| echo: false
#| results: asis
render_mean_diff_heatmap_tabs(
  rand_ilr_half = random_ilr_random_forest_pc_half,
  rand_ilr_max = random_ilr_random_forest_pc_max_lib_size,
  rod_half = rodriguez_random_forest_pc_half,
  rod_max = rodriguez_random_forest_pc_max_lib_size,
  bench_half = benchmark_random_forest_pc_half,
  bench_max = benchmark_random_forest_pc_max_lib_size,
  plot_metric = "misclassification_rate"
)
```
:::

# XGBoost
## ROC AUC

::: panel-tabset
```{r}
#| label: xgboost_roc_auc
#| echo: false
#| results: asis
render_mean_diff_heatmap_tabs(
  rand_ilr_half = random_ilr_xgboost_pc_half,
  rand_ilr_max = random_ilr_xgboost_pc_max_lib_size,
  rod_half = rodriguez_xgboost_pc_half,
  rod_max = rodriguez_xgboost_pc_max_lib_size,
  bench_half = benchmark_xgboost_pc_half,
  bench_max = benchmark_xgboost_pc_max_lib_size,
  plot_metric = "roc_auc"
)
```
:::

## Misclassification Rate
::: panel-tabset
```{r}
#| label: xgboost_misclas
#| echo: false
#| results: asis
render_mean_diff_heatmap_tabs(
  rand_ilr_half = random_ilr_xgboost_pc_half,
  rand_ilr_max = random_ilr_xgboost_pc_max_lib_size,
  rod_half = rodriguez_xgboost_pc_half,
  rod_max = rodriguez_xgboost_pc_max_lib_size,
  bench_half = benchmark_xgboost_pc_half,
  bench_max = benchmark_xgboost_pc_max_lib_size,
  plot_metric = "misclassification_rate"
)
```
:::
