---
title: "Results Summary: Lupus Dataset"
author: "Oktawia Miluch"
date: today
bibliography: ../readme_files/references.bib
link-citations: true
format:
  html:
    toc: true
    number-sections: true
    theme: cosmo
    fig-format: png
    code-fold: true
    embed-resources: true
execute:
  echo: true
  warning: false
  message: false
knitr:
  opts_chunk:
    fig.path: _figures/summary_lupus/
    fig.width: 18
    fig.height: 13.5
---

```{r}
#| label: setup
#| include: false
library(here)
library(purrr)
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(viridis)
library(kableExtra)

# Create/clean graphics directory, set fig.path
graphics_abs <- here::here("content", "_figures", "summary_lupus")
dir.create(graphics_abs, recursive = TRUE, showWarnings = FALSE)

# clean contents but keep folder
to_remove <- list.files(
  graphics_abs,
  all.files = TRUE,
  full.names = TRUE,
  no.. = TRUE
)
unlink(to_remove, recursive = TRUE, force = TRUE)

# ensure knitr writes to the selected path
knitr::opts_chunk$set(
  fig.path   = "_figures/summary_lupus/",
  fig.width  = 18,
  fig.height = 13.5
)
```

```{r}
#| label: data_read
#| include: false
# Source helper functions
source(file.path(here::here(), "R", "summary_agp_lupus.R"))
dir_results <- here::here("results", "lupus_results")
# Read in results
res <- read_results(dir_results)
```

In this script, 4 models are considered: 3 already used in the previous scripts, namely logistic regression with L1 penalty, Random Forest, and XGBoost. The fourth model is the sparse log-contrast classification model as implemented in the R package `trac` version 0.0.2 (see [viettr/trac](https://github.com/viettr/trac/tree/master) GitHub repository for implementation details). 

**Sparse log-contrast classification model with square hinge loss:**  The model can be described as follows:
$$
y = X \beta + \sigma \epsilon \quad \text{subject to} \quad C \beta = 0
$$

where $y$ is a binary outcome vector, $X$ is a predictor matrix, $C$ is a general constraint matrix,  $\beta$ is an unknown coefficient vector and  $\sigma$ is an unknown scale parameter. In this analysis the squared hinge loss is used with an $\ell_1$ penalty for the $\beta$ coefficients. Therefore, the optimization problem is defined as follows:

$$
\begin{aligned}
\min_{\beta \in \mathbb{R}^{d}}
\quad & \sum_{i=1}^{n} \, \ell\!\left(y_i\, x_i^{\top}\beta\right) \;+\; \lambda \lVert \beta \rVert_{1} \\
\text{subject to}\quad & C\beta = 0 ,
\end{aligned}
$$

where the $x_i$ are the rows of $X$, $\lVert \beta \rVert_{1}$ is the $\ell_1$ penalty and the loss function $\ell(\cdot)$ is defined as:

$$
\ell(r)=
\begin{cases}
(1-r)^2, & \text{if } r \le 1,\\[4pt]
0, & \text{if } r \ge 1.
\end{cases}
$$

Given this setup, the fitted model outputs a score for each observation, which can be thresholded to obtain class predictions. To obtain probabilities, an additional step is required (e.g., additional model transforming scores into probabilities). Consequently, Brier scores are not reported for the sparse log-contrast model. The sparse log contrast model is fitted only on the original data (not augmented) and its purpose is to provide an additional benchmark for the logistic regression model with L1 penalty.

# Logistic regression with L1 penalty
## Summary Table
```{r}
#| message: false
#| warning: false
summary_tbl_mean_perf(res$lasso, res$sparse_log_contrast)
```

## Boxplots of performance metrics
```{r}
#| label: boxplots_lasso
boxplot_metric(res$lasso, res$sparse_log_contrast, plot_metric = "roc_auc")
boxplot_metric(res$lasso, res$sparse_log_contrast, 
plot_metric = "misclassification_rate")
boxplot_metric(res$lasso, plot_metric = "brier_class")
```

## ROC Curves
```{r}
roc_curve_plot(res$lasso, res$sparse_log_contrast)
```

## Density Impact
```{r}
dir_results_density <- here::here("results", "lupus_density_results")
# Read in results - impact of density 
res_density <- read_results(dir_results_density)
boxplot_density(res_density$lasso, plot_metric = "roc_auc")
boxplot_density(res_density$lasso, plot_metric = "misclassification_rate")
```

## Interpretability
The purpose of the two subchapters below is to better understand which original features (or which contrasts) are most influential for a specific data augmentation strategy.

### Global view
The table below shows the average number of $\beta$ coefficients (average over $20$ train-test splits of the dataset) that were estimated to be non-zero in the logistic regression model with $L1$ penalty and the total number of estimated coefficients for different variants of the model. 

```{r}
#| label: lasso_non_zero_coefs_avg
tables_dir <- here::here("content", "_tables", "summary_lupus")
dir.create(tables_dir, recursive = TRUE, showWarnings = FALSE)

log_reg_tbl <- log_reg_summary_tbl(res)
log_reg_tbl_latex <- log_reg_summary_tbl(res, format = "latex")

kableExtra::save_kable(
  log_reg_tbl_latex,
  file = file.path(tables_dir, "lasso_non_zero_coefs_avg.tex")
)

log_reg_tbl
```

In the case of random ilr augmentation in p and the combination of Aitchison Mixup and augmentation in p, one can analyse how many times, on average among $20$ train-test splits, the $\beta$ coefficients that were estimated to be non-zero are associated with the standard or random ilr transformation. The table below provides such a summary:

```{r}
avg_random_tbl <- avg_random_vs_standard(res)
avg_random_tbl_latex <- avg_random_vs_standard(res, format = "latex")

kableExtra::save_kable(
  avg_random_tbl_latex,
  file = file.path(tables_dir, "avg_random_vs_standard.tex")
)

avg_random_tbl
```

The purpose of the section below is to gain a better understanding of the original features that contribute to standard ilr coordinates and, if applicable, to the random ilr coordinates that were selected by the model (have non-zero beta coefficients). The table below shows the original features that contribute to the numerator of the standard ILR coordinates that are selected in (almost) all 20 data splits. For example, in the model trained on the data augmented in p, the ILR coordinate associated with the standard ILR transformation that has the feature T4_em in the numerator is selected in all 20 data splits.

```{r}
# Get names of the features in the input data:
data_dir_lupus <- here::here("data", "data_lupus", "data_preproc")
data_lupus <- readRDS(file.path(data_dir_lupus, "data_lupus_prep.RDS"))
cov_names <- tibble(row_idx = 1:ncol(data_lupus$prop_pc), 
  covariate = data_lupus$prop_pc |> colnames()
) |>
mutate(covariate = str_remove(covariate, "g__"))
rm(data_lupus)

# Get contributors to the ilr coordinates that are selected by a models from all 20 test-train splits
# benchmark
benchmark_contr_tbl <- contr_standard_ilr_all_splits(res, dir_results =  dir_results, sel_aug_strategy = NA, model = "lasso")
# Aitchison Mixup
aitchison_contr_tbl <- contr_standard_ilr_all_splits(res, dir_results =  dir_results, sel_aug_strategy = "aitchison_mixup", model = "lasso")
# Augmentation in p
aug_p_contr_tbl <- contributors_all_splits(dir_results = dir_results, sel_aug_strategy = "aug_in_p", top_contr = 3, res = res, model = "lasso")
# Aitchison + augmentation in p
aitchison_aug_p_contr_tbl <- contributors_all_splits(dir_results = dir_results, sel_aug_strategy = "aitchison_aug_in_p", top_contr = 3, res = res, model = "lasso")

# Most influential contributors for standard ILR across 20 data splits
top_k_standard_tbl <- top_k_standard_ilr(
  tbls = list(
    benchmark_contr_tbl,
    aitchison_contr_tbl,
    aug_p_contr_tbl,
    aitchison_aug_p_contr_tbl
  ),
  top_n = c(6, 10, 5, 7),
  section_titles = c(
    "Benchmark",
    "Aitchison Mixup",
    "Augmentation in p",
    "Aitchison Mixup & Augmentation in p"
  ),
  cov_names = cov_names
)

top_k_standard_tbl_latex <- top_k_standard_ilr(
  tbls = list(
    benchmark_contr_tbl,
    aitchison_contr_tbl,
    aug_p_contr_tbl,
    aitchison_aug_p_contr_tbl
  ),
  top_n = c(6, 10, 5, 7),
  section_titles = c(
    "Benchmark",
    "Aitchison Mixup",
    "Augmentation in p",
    "Aitchison Mixup & Augmentation in p"
  ),
  cov_names = cov_names,
  format = "latex",
  latex_table_env = "tabularx"
)

kableExtra::save_kable(
  top_k_standard_tbl_latex,
  file = file.path(tables_dir, "top_k_standard_ilr.tex")
)

top_k_standard_tbl
```

The below table summarizes the original features that are most often in the top 3 contributors to the random ILR coordinates that get selected by the model ($\beta \neq  0$). For example, in the case of data augmented in p, feature Prolif is within top 3 contributors to the 31 random ILR coordinates that are selected by the models obtained from 20 data splits. 

```{r}
#| lable: random_ilr_contr_lasso
### Random ILR part
aug_random_tbl <- aug_p_contr_tbl |>
 unnest(random_ilr) |> 
 count(row_idx) |>
 arrange(desc(n)) |>
 left_join(cov_names, by = "row_idx") |>
 select(covariate, n) |>
 head(n = 5)

aitchison_random_tbl <- aitchison_aug_p_contr_tbl |>
 unnest(random_ilr) |> 
 count(row_idx) |>
 arrange(desc(n)) |>
 left_join(cov_names, by = "row_idx") |>
 select(covariate, n) |>
 head(n = 6)

# Table showing features that are most often among top 3 contributors
random_ilr_combined <- bind_rows(
  aug_random_tbl,
  aitchison_random_tbl
)

random_ilr_tbl <- random_ilr_combined |>
  knitr::kable(
    col.names = c("Original feature", "# of random ILR coordinates"),
    caption = "Summary of features that are most often within top 3 contributors to the selected (est. coefs non-zero) **random ILR** coordinates",
    escape = FALSE,
    align = c("l", "c")
  ) |>
  kableExtra::group_rows("Augmentation in p", 1, nrow(aug_random_tbl)) |>
  kableExtra::group_rows("Aitchison Mixup & Augmentation in p", nrow(aug_random_tbl) + 1, nrow(aug_random_tbl) + nrow(aitchison_random_tbl)) |>
  kableExtra::kable_styling(full_width = TRUE)

random_ilr_tbl_latex <- random_ilr_combined |>
  knitr::kable(
    col.names = c("Original feature", "# of random ILR coordinates"),
    caption = "Summary of features that are most often within top 3 contributors to the selected (est. coefs non-zero) **random ILR** coordinates",
    escape = FALSE,
    align = c("l", "c"),
    format = "latex"
  ) |>
  kableExtra::group_rows("Augmentation in p", 1, nrow(aug_random_tbl)) |>
  kableExtra::group_rows("Aitchison Mixup & Augmentation in p", nrow(aug_random_tbl) + 1, nrow(aug_random_tbl) + nrow(aitchison_random_tbl)) |>
  kableExtra::kable_styling(full_width = FALSE, latex_table_env = "tabularx")

kableExtra::save_kable(
  random_ilr_tbl_latex,
  file = file.path(tables_dir, "random_ilr_contr_lasso.tex")
)

random_ilr_tbl
```

### Local view

Below I focus only on one test-train split and try to understand which original features constitute the ilr coordinates that are selected by model. I analyse data split number 11 because for majority of augmentation strategies this split leads to the highest ROC AUC on the test data. 

#### Benchmark
```{r}
#| include: false
sel_coords <- benchmark_contr_tbl |> 
  filter(split == 11) |> 
  unnest(standard_ilr) |>  
  nrow()

```

For analysed data split, in the benchmark model `` `r sel_coords` `` standard ILR coordinates were selected.

```{r}
# Benchmark
standard_ilr_one_split(benchmark_contr_tbl, 11, cov_names = cov_names)

standard_ilr_one_split(
  benchmark_contr_tbl,
  11,
  cov_names = cov_names,
  format = "latex",
  output_dir = tables_dir,
  file_name = "standard_ilr_one_split_benchmark_11.tex"
)
```

#### Aitchison Mixup
```{r}
#| include: false
sel_coords <- aitchison_contr_tbl |> 
  filter(split == 11) |> 
  unnest(standard_ilr) |>  
  nrow()
```

For analysed data split, in the model trained on Aitchison Mixup augmented data `` `r sel_coords` `` standard ILR coordinates were selected.

```{r}
# Aitchison Mixup
standard_ilr_one_split(aitchison_contr_tbl, 11, cov_names = cov_names)
```

#### Augmentation in p

```{r}
#| include: false
sel_coords_st <- aug_p_contr_tbl |> 
  filter(split == 11) |> 
  unnest(standard_ilr) |>  
  nrow()

sel_coord_rand <- aug_p_contr_tbl |> 
  filter(split == 11)  |>
  unnest(random_ilr) |> 
  distinct(segment, ilr_coordinate) |> 
  nrow()
```

For analysed data split, in the model trained on data augmented in p `` `r sel_coords_st` `` standard ILR coordinates and `` `r sel_coord_rand` `` random ILR were selected.

```{r}
# Aug in p
## Random ilr part
heatmap_contr(aug_p_contr_tbl, sel_split = 11, cov_names = cov_names)

## Standard ilr part
standard_ilr_one_split(aug_p_contr_tbl, 11, cov_names = cov_names)
standard_ilr_one_split(
  aug_p_contr_tbl,
  11,
  cov_names = cov_names,
  format = "latex",
  output_dir = tables_dir,
  file_name = "standard_ilr_one_split_aug_p_11.tex"
)
```

The plot below shows the estimated values of the $\beta$ coefficients in the model fitted on the data  augmented in p, corresponding to the analysed data split. The coefficients are grouped by the ILR base used to create the feature with that coefficient. As can be seen, the largest coefficients are assigned to the coordinates originating from the standard ILR transformation.

```{r}
#\ label: lollipop_plot_lasso
# Lolipop plot with coefficients 
df_plot <- res$lasso |> 
  filter(aug_strategy == "aug_in_p", split_seed == 11) |> 
  select(-c(aug_factor:lambda_1se)) |> 
  unnest(beta_1se) |> 
  filter(estimate != 0, term != "(Intercept)") |>
  mutate(
      base_id = if_else(str_detect(term, "_\\d+$"),
                        str_extract(term, "(?<=_)\\d+$"),
                        "standard"),
      base = if_else(base_id == "standard",
                    "Standard ILR",
                    paste("Random ILR base", base_id)),
      coord = str_remove(term, "_\\d+$") |> str_remove("^V") |> as.integer(),
      coord_label = paste0("Coordinate ", coord)
    )

base_levels <- c(
  "Standard ILR",
  paste("Random ILR base", sort(unique(df_plot$base_id[df_plot$base_id != "standard"])))
)

df_plot <- df_plot |>
  mutate(base = factor(base, levels = base_levels)) |>
  arrange(base, desc(abs(estimate))) |>
  mutate(y = row_number())

base_bounds <- df_plot |>
  group_by(base) |>
  summarise(
    y_min = min(y),
    y_max = max(y),
    y_mid = (y_min + y_max) / 2,
    .groups = "drop"
  ) |>
  arrange(base)

sep_lines <- base_bounds |>
  mutate(y_sep = y_max + 0.5) |>
  slice_head(n = 2)

x_lab_inside <- 1
x_min <- min(df_plot$estimate, 0, x_lab_inside)
x_max <- max(df_plot$estimate, x_lab_inside)

ggplot(df_plot, aes(x = estimate, y = y, color = base)) +
  geom_segment(aes(x = 0, xend = estimate, yend = y), linewidth = 1.1) +
  geom_point(size = 3.2) +
  geom_hline(data = sep_lines, aes(yintercept = y_sep), color = "grey70") +
  geom_text(
    data = base_bounds,
    aes(x = x_lab_inside, y = y_mid, label = base),
    hjust = 0,
    fontface = "bold",
    size = 4.5,
    color = "grey20"
  ) +
  scale_y_continuous(breaks = df_plot$y, labels = df_plot$coord_label) +
  scale_color_viridis_d(option = "D", end = 0.9, guide = "none")  +
  labs(
    x = expression(paste("Estimates of ", beta, " coefficients")),
    y = NULL,
    title = "Coefficients of L1-penalized logistic regression by ILR basis",
    subtitle = "Data split: 11"
  ) +
  coord_cartesian(xlim = c(x_min, x_max)) +
  theme_bw()
```

#### Aitchison Mixup + Augmentation in p

```{r}
#| include: false
sel_coords_st <- aitchison_aug_p_contr_tbl |> 
  filter(split == 11) |> 
  unnest(standard_ilr) |>  
  nrow()

sel_coord_rand <- aitchison_aug_p_contr_tbl |> 
  filter(split == 11)  |>
  unnest(random_ilr) |> 
  distinct(segment, ilr_coordinate) |> 
  nrow()
```

For analysed data split, in the model trained on data augmented by Aitchison Mixup and augmented in p, `` `r sel_coords_st` `` standard ILR coordinates and `` `r sel_coord_rand` `` random ilr coordinates were selected.


```{r}
# Aitchison + aug in p
heatmap_contr(aitchison_aug_p_contr_tbl, sel_split = 11, cov_names = cov_names)

standard_ilr_one_split(aitchison_aug_p_contr_tbl, 11, cov_names = cov_names)
```

# Random Forest
## Summary Table
```{r}
#| message: false
#| warning: false
summary_tbl_mean_perf(res$RF)
```

## Boxplots of performance metrics
```{r}
boxplot_metric(res$RF, plot_metric = "roc_auc")
boxplot_metric(res$RF, plot_metric = "misclassification_rate")
boxplot_metric(res$RF, plot_metric = "brier_class")
```

## ROC Curves
```{r}
roc_curve_plot(res$RF)
```

## Density Impact
```{r}
boxplot_density(res_density$RF, plot_metric = "roc_auc")
boxplot_density(res_density$RF, plot_metric = "misclassification_rate")
```

## Feature Importance
Feature importance is computed on the test data. For each model trained on one data split, 5 ilr coordinates with the higest variable importance scores are identified. Then, for those coordinates, 3 original features that contributed most to them are identified. 
Below table shows original features that contribute most often to standard ILR coordiantes that achieve the highest VIMP scores. For example, feature T4_em contributes in numerator of the standard ILR coordinate that is selected by all models in nearly all data splits.

```{r}
# Get contributors to the ilr coordinates that achieve the highest VIMP in a model across all 20 test-train splits
# for each model (trained on 1 data split) select num_top_feat ilr coordinates that in this model achieved the highest VIMP --> decode original features that contributed most to this ilr coordinates
# benchmark
benchmark_contr_tbl <- contr_standard_ilr_all_splits(res, dir_results =  dir_results, sel_aug_strategy = NA, model = "RF", num_top_feat = 5)
# Aitchison Mixup
aitchison_contr_tbl <- contr_standard_ilr_all_splits(res, dir_results =  dir_results, sel_aug_strategy = "aitchison_mixup", model = "RF", num_top_feat = 5)
# Augmentation in p
aug_p_contr_tbl <- contributors_all_splits(dir_results = dir_results, sel_aug_strategy = "aug_in_p", top_contr = 3, res = res, model = "RF", num_top_feat = 5)
# Aitchison + augmentation in p
aitchison_aug_p_contr_tbl <- contributors_all_splits(dir_results = dir_results, sel_aug_strategy = "aitchison_aug_in_p", top_contr = 3, res = res, model = "RF", num_top_feat = 5)

# Most influential contributors for standard ILR across 20 data splits
top_k_standard_tbl <- top_k_standard_ilr(
  tbls = list(
    benchmark_contr_tbl,
    aitchison_contr_tbl,
    aug_p_contr_tbl,
    aitchison_aug_p_contr_tbl
  ),
  top_n = rep(3, 4),
  section_titles = c(
    "Benchmark",
    "Aitchison Mixup",
    "Augmentation in p",
    "Aitchison Mixup & Augmentation in p"
  ),
  cov_names = cov_names
)

top_k_standard_tbl_latex <- top_k_standard_ilr(
  tbls = list(
    benchmark_contr_tbl,
    aitchison_contr_tbl,
    aug_p_contr_tbl,
    aitchison_aug_p_contr_tbl
  ),
  top_n = rep(3, 4),
  section_titles = c(
    "Benchmark",
    "Aitchison Mixup",
    "Augmentation in p",
    "Aitchison Mixup & Augmentation in p"
  ),
  cov_names = cov_names,
  format = "latex",
  latex_table_env = "tabularx"
)

kableExtra::save_kable(
  top_k_standard_tbl_latex,
  file = file.path(tables_dir, "top_k_standard_ilr_RF.tex")
)

top_k_standard_tbl

```

The below table summarizes the original features that are most often in the top 3 contributors to the random ILR coordinates that achieve the highest VIMP. Below frequencies are significantly smaller than in the case or logistic regression with L1 penalty. This is the consequence of the fact that I filter out all ILR coordinates that have negative importance so we might end up with no features from a given segment (standard or random ILR).

```{r}
#| label: random_ilr_contr_rf
### Random ILR part
aug_random_tbl <- aug_p_contr_tbl |>
 unnest(random_ilr) |> 
 count(row_idx) |>
 arrange(desc(n)) |>
 left_join(cov_names, by = "row_idx") |>
 select(covariate, n) |>
 head(n = 5)

aitchison_random_tbl <- aitchison_aug_p_contr_tbl |>
 unnest(random_ilr) |> 
 count(row_idx) |>
 arrange(desc(n)) |>
 left_join(cov_names, by = "row_idx") |>
 select(covariate, n) |>
 head(n = 4)

# Table showing features that are most often among top 3 contributors
random_ilr_combined <- bind_rows(
 aug_random_tbl,
 aitchison_random_tbl
)

random_ilr_tbl <- random_ilr_combined |>
 knitr::kable(
  col.names = c("Original feature", "# of random ILR coordinates"),
  caption = "Summary of features that are most often within top 3 contributors to the **random ILR** coordinates with the highest VIMP",
  escape = FALSE,
  align = c("l", "c")
 ) |>
 kableExtra::group_rows("Augmentation in p", 1, nrow(aug_random_tbl)) |>
 kableExtra::group_rows("Aitchison Mixup & Augmentation in p", nrow(aug_random_tbl) + 1, nrow(aug_random_tbl) + nrow(aitchison_random_tbl)) |>
 kableExtra::kable_styling(full_width = TRUE)

random_ilr_tbl_latex <- random_ilr_combined |>
 knitr::kable(
  col.names = c("Original feature", "# of random ILR coordinates"),
  caption = "Summary of features that are most often within top 3 contributors to the **random ILR** coordinates with the highest VIMP",
  escape = FALSE,
  align = c("l", "c"),
  format = "latex"
 ) |>
 kableExtra::group_rows("Augmentation in p", 1, nrow(aug_random_tbl)) |>
 kableExtra::group_rows("Aitchison Mixup & Augmentation in p", nrow(aug_random_tbl) + 1, nrow(aug_random_tbl) + nrow(aitchison_random_tbl)) |>
 kableExtra::kable_styling(full_width = FALSE, latex_table_env = "tabularx")

kableExtra::save_kable(
  random_ilr_tbl_latex,
  file = file.path(tables_dir, "random_ilr_contr_rf.tex")
)

random_ilr_tbl
```

### Local view

Below I focus only on one test-train split and try to understand which original features constitute the ilr coordinates that obtain high VIMP. I analyse data split number 11 because for all augmentation strategies this split leads to the highest ROC AUC on the test data. 

#### Benchmark
```{r}
#| include: false
sel_coords <- benchmark_contr_tbl |> 
  filter(split == 11) |> 
  unnest(standard_ilr) |>  
  nrow()
```

For analysed data split, in the benchmark model `` `r sel_coords` `` standard ILR coordinates were selected.

```{r}
# Benchmark
standard_ilr_one_split(benchmark_contr_tbl, 11, cov_names = cov_names)
```

#### Aitchison Mixup
```{r}
#| include: false
sel_coords <- aitchison_contr_tbl |> 
  filter(split == 11) |> 
  unnest(standard_ilr) |>  
  nrow()
```

For analysed data split, in the model trained on Aitchison Mixup augmented data `` `r sel_coords` `` standard ILR coordinates were selected.

```{r}
# Aitchison Mixup
standard_ilr_one_split(aitchison_contr_tbl, 11, cov_names = cov_names)
```

#### Augmentation in p

```{r}
#| include: false
sel_coords_st <- aug_p_contr_tbl |> 
  filter(split == 11) |> 
  unnest(standard_ilr) |>  
  nrow()

sel_coord_rand <- aug_p_contr_tbl |> 
  filter(split == 11)  |>
  unnest(random_ilr) |> 
  summarise(val = if (n() == 0) 0 else n_distinct(segment, ilr_coordinate)) |>
  pull(val)
```

For analysed data split, in the model trained on data augmented in p `` `r sel_coords_st` `` standard ILR coordinates and `` `r sel_coord_rand` `` random ILR coordinates were selected.

```{r}
# Aug in p
## Random ilr part
heatmap_contr(aug_p_contr_tbl, sel_split = 11, cov_names = cov_names)

## Standard ilr part
standard_ilr_one_split(aug_p_contr_tbl, 11, cov_names = cov_names)
```

#### Aitchison Mixup + Augmentation in p

```{r}
#| include: false
sel_coords_st <- aitchison_aug_p_contr_tbl |> 
  filter(split == 11) |> 
  unnest(standard_ilr) |>  
  nrow()

sel_coord_rand <- aitchison_aug_p_contr_tbl |>
  filter(split == 11) |>
  unnest(random_ilr) |>
  summarise(val = if (n() == 0) 0 else n_distinct(segment, ilr_coordinate)) |>
  pull(val)
```

For analysed data split, in the model trained on data augmented by Aitchison Mixup and augmented in p `` `r sel_coords_st` `` standard ILR coordinates and `` `r sel_coord_rand` `` random ILR coordinates were selected.


```{r}
standard_ilr_one_split(aitchison_aug_p_contr_tbl, 11, cov_names = cov_names)
```


# XGBoost
## Summary Table
```{r}
#| message: false
#| warning: false
summary_tbl_mean_perf(res$XGB)
```

## Boxplots of performance metrics
```{r}
boxplot_metric(res$XGB, plot_metric = "roc_auc")
boxplot_metric(res$XGB, plot_metric = "misclassification_rate")
boxplot_metric(res$XGB, plot_metric = "brier_class")
```

## ROC Curves
```{r}
roc_curve_plot(res$XGB)
```

## Density Impact
```{r}
boxplot_density(res_density$XGB, plot_metric = "roc_auc")
boxplot_density(res_density$XGB, plot_metric = "misclassification_rate")
```

## Feature Importance
Feature importance is computed on the test data. Below table shows original features that contribute most often to standard ILR coordiantes that achieve the highest VIMP scores. For example, feature T4_em contributes in numerator of the standard ILR coordinate that is selected by all models in nearly all data splits.

```{r}
# Get contributors to the ilr coordinates that achieve the highest VIMP in a model across all 20 test-train splits
# for each model (trained on 1 data split) select num_top_feat ilr coordinates that in this model achieved the highest VIMP --> decode original features that contributed most to this ilr coordinates
# benchmark
benchmark_contr_tbl <- contr_standard_ilr_all_splits(res, dir_results =  dir_results, sel_aug_strategy = NA, model = "XGB", num_top_feat = 5)
# Aitchison Mixup
aitchison_contr_tbl <- contr_standard_ilr_all_splits(res, dir_results =  dir_results, sel_aug_strategy = "aitchison_mixup", model = "XGB", num_top_feat = 5)
# Augmentation in p
aug_p_contr_tbl <- contributors_all_splits(dir_results = dir_results, sel_aug_strategy = "aug_in_p", top_contr = 3, res = res, model = "XGB", num_top_feat = 5)

# Aitchison + augmentation in p
aitchison_aug_p_contr_tbl <- contributors_all_splits(dir_results = dir_results, sel_aug_strategy = "aitchison_aug_in_p", top_contr = 3, res = res, model = "XGB", num_top_feat = 5)

# Most influential contributors for standard ILR across 20 data splits
top_k_standard_ilr(
  tbls = list(
    benchmark_contr_tbl,
    aitchison_contr_tbl,
    aug_p_contr_tbl,
    aitchison_aug_p_contr_tbl
  ),
  top_n = rep(4, 4),
  section_titles = c(
    "Benchmark",
    "Aitchison Mixup",
    "Augmentation in p",
    "Aitchison Mixup & Augmentation in p"
  ),
  cov_names = cov_names
)
```

The below table summarizes the original features that are most often in the top 3 contributors to the random ILR coordinates that achieve the highest VIMP. Below frequencies are significantly smaller than in the case or logistic regression with L1 penalty. This is the consequence of the fact that I filter out all ILR coordinates that have negative importance so we might end up with no features from a given segment (standard or random ILR).

```{r}
### Random ILR part
aug_random_tbl <- aug_p_contr_tbl |>
 unnest(random_ilr) |> 
 count(row_idx) |>
 arrange(desc(n)) |>
 left_join(cov_names, by = "row_idx") |>
 select(covariate, n) |>
 head(n = 5)

aitchison_random_tbl <- aitchison_aug_p_contr_tbl |>
 unnest(random_ilr) |> 
 count(row_idx) |>
 arrange(desc(n)) |>
 left_join(cov_names, by = "row_idx") |>
 select(covariate, n) |>
 head(n = 4)

# Table showing features that are most often among top 3 contributors
bind_rows(
 aug_random_tbl,
 aitchison_random_tbl
) |>
 knitr::kable(
  col.names = c("Original feature", "# of random ILR coordinates"),
  caption = "Summary of features that are most often within top 3 contributors to the **random ILR** coordinates with the highest VIMP",
  escape = FALSE,
  align = c("l", "c", "c")
 ) |>
 kableExtra::group_rows("Augmentation in p", 1, nrow(aug_random_tbl)) |>
 kableExtra::group_rows("Aitchison Mixup & Augmentation in p", nrow(aug_random_tbl) + 1, nrow(aug_random_tbl) + nrow(aitchison_random_tbl)) |>
 kableExtra::kable_styling(full_width = TRUE)
```

### Local view

Below I focus only on one test-train split and try to understand which original features constitute the ilr coordinates that obtain the highest VIMP. I analyse data split number 11 because for all augmentation strategies this split leads to the highest ROC AUC on the test data. 

#### Benchmark
```{r}
#| include: false
sel_coords <- benchmark_contr_tbl |> 
  filter(split == 11) |> 
  unnest(standard_ilr) |>  
  nrow()
```

For analysed data split, in the benchmark model `` `r sel_coords` `` standard ILR coordinates were selected.

```{r}
# Benchmark
standard_ilr_one_split(benchmark_contr_tbl, 11, cov_names = cov_names)
```

#### Aitchison Mixup
```{r}
#| include: false
sel_coords <- aitchison_contr_tbl |> 
  filter(split == 11) |> 
  unnest(standard_ilr) |>  
  nrow()
```

For analysed data split, in the model trained on Aitchison Mixup augmented data `` `r sel_coords` `` standard ILR coordinates were selected.

```{r}
# Aitchison Mixup
standard_ilr_one_split(aitchison_contr_tbl, 11, cov_names = cov_names)
```

#### Augmentation in p

```{r}
#| include: false
sel_coords_st <- aug_p_contr_tbl |> 
  filter(split == 11) |> 
  unnest(standard_ilr) |>  
  nrow()

sel_coord_rand <- aug_p_contr_tbl |> 
  filter(split == 11)  |>
  unnest(random_ilr) |> 
  summarise(val = if (n() == 0) 0 else n_distinct(segment, ilr_coordinate)) |>
  pull(val)
```

For analysed data split, in the model trained on data augmented in p `` `r sel_coords_st` `` standard ILR coordinates and `` `r sel_coord_rand` `` random ILR coordinates were selected.

```{r}
# Aug in p
## Standard ilr part
standard_ilr_one_split(aug_p_contr_tbl, 11, cov_names = cov_names)
```

#### Aitchison Mixup + Augmentation in p

```{r}
#| include: false
sel_coords_st <- aitchison_aug_p_contr_tbl |> 
  filter(split == 11) |> 
  unnest(standard_ilr) |>  
  nrow()

sel_coord_rand <- aitchison_aug_p_contr_tbl |>
  filter(split == 11) |>
  unnest(random_ilr) |>
  summarise(val = if (n() == 0) 0 else n_distinct(segment, ilr_coordinate)) |>
  pull(val)
```

For analysed data split, in the model trained on data augmented by Aitchison Mixup and augmented in p `` `r sel_coords_st` `` standard ILR coordinates and `` `r sel_coord_rand` `` random ILR coordinates were selected.

```{r}
standard_ilr_one_split(aitchison_aug_p_contr_tbl, 11, cov_names = cov_names)
```



