---
title: "Results Summary: Lupus Dataset"
author: "Oktawia Miluch"
date: today
bibliography: ../readme_files/references.bib
link-citations: true
format:
  html:
    toc: true
    number-sections: true
    theme: cosmo
    fig-format: png
    code-fold: true
    embed-resources: true
execute:
  echo: true
  warning: false
  message: false
knitr:
  opts_chunk:
    fig.path: _figures/summary_lupus/
    fig.width: 18
    fig.height: 13.5
---

```{r}
#| label: setup
#| include: false
library(here)
library(purrr)
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(viridis)
library(kableExtra)

# Create/clean graphics directory, set fig.path
graphics_abs <- here::here("content", "_figures", "summary_lupus")
dir.create(graphics_abs, recursive = TRUE, showWarnings = FALSE)

# clean contents but keep folder
to_remove <- list.files(
  graphics_abs,
  all.files = TRUE,
  full.names = TRUE,
  no.. = TRUE
)
unlink(to_remove, recursive = TRUE, force = TRUE)

# ensure knitr writes to the selected path
knitr::opts_chunk$set(
  fig.path   = "_figures/summary_lupus/",
  fig.width  = 18,
  fig.height = 13.5
)
```

```{r}
#| label: data_read
#| include: false
# Source helper functions
source(file.path(here::here(), "R", "summary_agp_lupus.R"))
dir_results <- here::here("results", "lupus_results")
# Read in results
res <- read_results(dir_results)
```

In this script 4 models are considered: 3 already used in the previous scripts, namely: logistic regression with L1 penalty, Random Forest and XGBoost. The fourth used model is the sparse log-contrast classification model as implemented in the R package `trac` version 0.0.2 (see [viettr/trac](https://github.com/viettr/trac/tree/master) GitHub repository for the implementation details). 

**Sparse log-contrast classification model with square hinge loss:**  The model can be described as follows:
$$
y = X \beta + \sigma \epsilon \quad \text{subject to} \quad C \beta = 0
$$

where $y$ is a binary outcome vector, $X$ is a predictor matrix, $C$ is a general constraint matrix,  $\beta$ is an unknown coefficient vector and  $\sigma$ is an unknown scale parameter. In this analysis the squared hinge loss is used with an $\ell_1$ penalty for the $\beta$ coefficients. Therefore, the optimization problem is defined as follows:

$$
\begin{aligned}
\min_{\beta \in \mathbb{R}^{d}}
\quad & \sum_{i=1}^{n} \, \ell\!\left(y_i\, x_i^{\top}\beta\right) \;+\; \lambda \lVert \beta \rVert_{1} \\
\text{subject to}\quad & C\beta = 0 ,
\end{aligned}
$$

where the $x_i$ are the rows of $X$, $\lVert \beta \rVert_{1}$ is the $\ell_1$ penalty and the loss function $\ell(\cdot)$ is defined as:

$$
\ell(r)=
\begin{cases}
(1-r)^2, & \text{if } r \le 1,\\[4pt]
0, & \text{if } r \ge 1.
\end{cases}
$$

Given this setup, the fitted model outputs a score for each observation, which can be thresholded to obtain class predictions. To obtain probabilities, an additional step is required (e.g., additional model transforming scores into probabilities). Consequently, Brier scores are not reported for the sparse log-contrast model. The sparse log contrast modle is fitted only on the original data (not augmented) and its purpose is to provide additional benchmark for the logistic regression model with L1 penalty.

# Logistic regression with L1 penalty
## Boxplots of performance metrics
```{r}
boxplot_metric(res$lasso, res$sparse_log_contrast, plot_metric = "roc_auc")
boxplot_metric(res$lasso, res$sparse_log_contrast, 
plot_metric = "misclassification_rate")
boxplot_metric(res$lasso, plot_metric = "brier_class")
```

## ROC Curves
```{r}
roc_curve_plot(res$lasso, res$sparse_log_contrast)
```

# Random Forest
## Boxplots of performance metrics
```{r}
boxplot_metric(res$RF, plot_metric = "roc_auc")
boxplot_metric(res$RF, plot_metric = "misclassification_rate")
boxplot_metric(res$RF, plot_metric = "brier_class")
```

## ROC Curves
```{r}
roc_curve_plot(res$RF)
```

# XGBoost
## Boxplots of performance metrics
```{r}
boxplot_metric(res$XGB, plot_metric = "roc_auc")
boxplot_metric(res$XGB, plot_metric = "misclassification_rate")
boxplot_metric(res$XGB, plot_metric = "brier_class")
```

## ROC Curves
```{r}
roc_curve_plot(res$XGB)
```